---
title: "OCR with Tesseract"
author: "Paul Bradshaw"
date: "20/02/2020"
output: html_document
---

# Trying out Tesseract for OCR on PDFs

Install Tesseract. [Vignette here](https://cran.r-project.org/web/packages/tesseract/vignettes/intro.html).

```{r}
install.packages("tesseract")
library(tesseract)
```

## Converting PDF to image

[From the vignette](https://cran.r-project.org/web/packages/tesseract/vignettes/intro.html#read_from_pdf_files):

> "If your images are stored in PDF files they first need to be converted to a proper image format. We can do this in R using the pdf_convert function from the pdftools package. Use a high DPI to keep quality of the image."

We use the [annual accounts from Victoria Beckham Ltd](https://beta.companieshouse.gov.uk/company/06517802/filing-history).

```{r}
pngfile <- pdftools::pdf_convert('victoriabeckhamltd2019.pdf', dpi = 600)
```

## Reading from the image of the PDF

Then see what text the OCR comes up with.

```{r}
text <- tesseract::ocr(pngfile)
fulltext <- cat(text)
```

